---
title: "Lab 7: Inference for Categorical DataR"
author: "Daniel Deneau"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1
These percentages appear to be sample statistics derived from the data sample provided WIN-Gallup

### Exercise 2

We must assume that the sampling method was random, and each sample was done independently. This does seem reasonable to do with  polls.

### Exercise 3

Each row in table 6 corresponds to a different country. Each row of atheism. Each row of the atheism table corresponds to an individual response by someone to the question of whether or not they are atheist.

```{r e3}
download.file("http://www.openintro.org/stat/data/atheism.RData", destfile = "atheism.RData")
load("atheism.RData")
atheism
```


### Exercise 4

The calculated proportion is:
0.05

The table shows the percentage of convinced atheists in the US to be 5%. so they do agree.
```{r e4}
us12 <- subset(atheism, nationality == "United States" & year == "2012")

us12_ath <- subset(us12, response=='atheist')
count(us12_ath) / count(us12)


```


### Exercise 5

First the samples must be random and independent, which they are since the poll was done randomly. Second, the success-failure criteria must be met:
```{r e5}
p = count(us12_ath)/count(us12)
n = count(us12)
n * p
n * (1 - p)

```
Which they are.

### Exercise 6
```{r e6}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")

```
The Margin of error is: 0.01 or 1%
```{r e6_2}
(1.95 * 0.0069)
```

### Exercise 7

For both Italy and Germany, the sampling is assumed to be random and independent.
The success-failure criteria for both countries are met  as well.
```{r e7}
it12 <- subset(atheism, nationality == "Italy" & year == "2012")
ge12 <- subset(atheism, nationality == "Germany" & year == "2012")
it_ath = subset(it12, response=="atheist")
ge_ath = subset(ge12, response=="atheist")
p_i = count(it_ath)/count(it12)
p_g = count(ge_ath)/count(ge12)
p_i * count(it12)
(1 - p_i) * count(it12)

p_g * count(ge12)
(1 - p_g) * count(ge12)
```

```{r e7 inference}
inference(it12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
inference(ge12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```
The margin of error for Italy is: 0.0086 or 0.86%
```{r e7-italy}
0.0086 * 1.95
```
The margin of error for Germany is: 0.0159 or 1.59%
```{r e7-germany}
0.0159 * 1.95
```


### Exercise 8
As the population proportion increases, the margin of error increases exponentially up to 50%. Once above 0.5, the margin of error decreases exponentially as the population proportion increases
```{r e8}
n <- 1000
p <- seq(0, 1, 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
plot(me ~ p, ylab = "Margin of Error", xlab = "Population Proportion")
```

### Exercise 9

The distribution is normal with a slight left skew and average spread.
The distribution is centered at the 0.1, with a mean 0f 0.09991. The interquartile range is .10577 - .09327 = 0.0125.
The range is .13558 - 0.06923 = 0.06635
```{r e9}
p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  p_hats[i] <- sum(samp == "atheist")/n
}

hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
summary(p_hats)

```

### Exercise 10
As n increases, the spread and standard deviation decreases and the mode becomes more precise.
As p increases, the distribution remains about the same with a slightly different center.
```{r e10}

p <- 0.1
n <- 1040
p_hats <- rep(0, 5000)

p_2 <- 0.1
n_2 <- 400
p_hats2 <- rep(0, 5000)

p_3 <- 0.02
n_3 <- 1040
p_hats3 <- rep(0, 5000)

p_4 <- 0.02
n_4 <- 400
p_hats4 <- rep(0, 5000)


for(i in 1:5000){
  samp <- sample(c("atheist", "non_atheist"), n, replace = TRUE, prob = c(p, 1-p))
  samp <- sample(c("atheist", "non_atheist"), n_2, replace = TRUE, prob = c(p_2, 1-p_2))
  samp <- sample(c("atheist", "non_atheist"), n_3, replace = TRUE, prob = c(p_3, 1-p_3))
  samp <- sample(c("atheist", "non_atheist"), n_4, replace = TRUE, prob = c(p_4, 1-p_4))
  p_hats[i] <- sum(samp == "atheist")/n
  p_hats2[i] <- sum(samp == "atheist")/n_2
  p_hats3[i] <- sum(samp == "atheist")/n_3
  p_hats4[i] <- sum(samp == "atheist")/n_4
}

par(mfrow = c(2, 2))
hist(p_hats, main = "p = 0.1, n = 1040", xlim = c(0, 0.18))
hist(p_hats2, main = "p = 0.1, n = 400", xlim = c(0, 0.18))
hist(p_hats3, main = "p = 0.2, n = 1040", xlim = c(0, 0.18))
hist(p_hats4, main = "p = 0.2, n = 400", xlim = c(0, 0.18))

```

### Exercise 11
Yes, because the distributions appear to be close to normal.

### OYO 1
a) Null Hypothesis: Spain has seen no change in its atheism index between 2005 and 2012
   Alt Hypothesis: Spain has seen some change in its atheism index between 2005 and 2012.
   The poll was random and independent, and the success-failure criteria are me:
   
```{r o1-sf}
sp12 <- subset(atheism, nationality == "Spain" & year == "2012")

sp12_ath = subset(sp12, response=="atheist")

sp05 <- subset(atheism, nationality == "Spain" & year == "2005")

sp05_ath = subset(sp05, response=="atheist")

p_s = count(sp12_ath)/count(sp12)

p_s * count(sp12)
(1 - p_s) * count(sp12)

p_s05 = count(sp05_ath)/count(sp05)

p_s05 * count(sp05)
(1 - p_s05) * count(sp05)

```
 The confidence levels do overlap quite a bit, therefore we can not reject the null hypothesis. There is not convincing evidence that Spain has seen a change.
```{r o1-inf}
inference(sp12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
inference(sp05$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```
 b) Null Hypothesis: The United States has seen no change in its atheism index between 2005 and 2012
   Alt Hypothesis: The United States has seen some change in its atheism index between 2005 and 2012.
   The poll was random and independent, and the success-failure criteria are me:
   
```{r o1-us}
us12 <- subset(atheism, nationality == "United States" & year == "2012")

us12_ath = subset(us12, response=="atheist")

us05 <- subset(atheism, nationality == "United States" & year == "2005")

us05_ath = subset(us05, response=="atheist")

p_u = count(sp12_ath)/count(sp12)

p_u * count(us12)
(1 - p_u) * count(us12)

p_u05 = count(us05_ath)/count(us05)

p_u05 * count(us05)
(1 - p_u) * count(sp05)

```
 The confidence levels do not overlap quite at all, therefore we can reject the null hypothesis. There is convincing evidence that the United States has seen a change.
```{r o1-inf-us}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
inference(us05$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```

### OYO 2
A type 1 error would be rejecting the null hypothesis that there is no change, when there is.
With 39 countries, I would expect to detect a change in 2 countries by chance, at a signifiance level of 0.05.

```{r o2}
0.05 * 39
```

### OYO 3
We would have to calculate the sample size from the standard error and population proportion, which is assumed to be 0.5.

```{r o3}
p = 0.5
m_error = 0.01
z = 1.95

standard_error = m_error/z

n = (p*p) / standard_error^2
n

```
So n must be at least n = 9506.25
n = 9507