---
title: "Lab 9: Linear Regression"
author: "Daniel Deneau"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
download.file("http://www.openintro.org/stat/data/mlb11.RData", destfile = "mlb11.RData")
load("mlb11.RData")
```

### Exercise 1

I would use a scatterplot.
The relationship does look linear, but there are some clouds of points with high runs and lower values of at_bats making the trend a bit weaker.
Overall, I would be comfortable using this model to predict the number of runs.

```{r code-e1}

plot(mlb11$at_bats, mlb11$runs)
cor(mlb11$runs, mlb11$at_bats)
mlb11
```

### Exercise 2
The plot shows a positive linear trend, that is strong with R = 0.61. We would see a stronger correlation if it were not for a cloud of observations towards the left side. The general trend is that the number of runs increases with as at_bat increases.
```{r code-e2}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
plot_ss(x = mlb11$at_bats, y = mlb11$runs, showSquares = TRUE)
```
### Exercise 3
The smallest sum of squares was 123721.9
```{r code-e3}
plot_ss(x = mlb11$at_bats, y = mlb11$runs)
```

### Exercise 4
y_hat = 415.2389 + 1.8345*homeruns
The slope tells us that the number of runs increases by 1.835 for each increase in the home runs.
```{r code-e4}
m1 <- lm(runs ~ at_bats, data = mlb11)
summary(m1)

m2 <- lm(runs ~ homeruns, data = mlb11)
summary(m2)

```

### Exercise 5
Using the equation, a team with 5,578 at-bats would have a predicted number of runs of: 728 runs.
```{r code-e51}
y <- -2789.2429 + (0.6305 * 5578)
y
```

Using the plotted pair (5579, 713), we see that 728 is an overstimate with the residual being: 728-713 = 15
```{r code-e5}
plot(mlb11$runs ~ mlb11$at_bats)
abline(m1)
data.frame(mlb11$runs, mlb11$at_bats)
728 - 713
```

### Excercise 6
There are no obvious patterns, so the relationship could be modeled linearly. We would need to perform a hypothesis test to determine whether the linearity is statistically significant or due to chance. 
```{r code-e6}
plot(m1$residuals ~ mlb11$at_bats)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
```

```{r code e7e8}
hist(m1$residuals)
qqnorm(m1$residuals)
qqline(m1$residuals)  # adds diagonal line to the normal prob plot
```
### Exercise 7
The histogram is nearly normal even though the shape is not perfect, and the normal probability plot follows a linear trend closely. So the residual conditions appear to be met.

### Exercise 8
The variability in both plots is not very high at all, so the constant variability condition appears to be met.

### OY0 1
I would like to see if players with less strikeouts have higher run counts. At a glance, there does appear to be a negative linear relationship.
```{r code-oyo1}
m3 <- lm(runs ~ strikeouts, data = mlb11)
plot(mlb11$runs ~ mlb11$strikeouts)
abline(m3)
```

### OYO 2
Only 16.9% of the variaility in runs is explained by strikeouts, compared to 37.3% for at-bats. So my model does not nearly predict runs at well.
```{r code-oyo2}
cor(mlb11$runs, mlb11$at_bats)
cor(mlb11$runs, mlb11$strikeouts)
summary(m1)
summary(m3)
```

### OYO 3
Batting average appears to best predict runs, with a correlation coefficient of 0.81
```{r code-oyo3}
m4 <- lm(runs ~ bat_avg, data = mlb11)
plot(mlb11$runs ~ mlb11$bat_avg)
cor(mlb11$runs, mlb11$bat_avg)
abline(m4)

hist(m4$residuals)
qqnorm(m4$residuals)
qqline(m4$residuals)

summary(m4)

```
### OYO 4
All of these variables do seem to be really good predictors for runs, with high correlation coefficients and strong linear trends in their plots. The best predictor is on-base plus slugging. This metric seems to take in account a lot of factors that makes high performing players have a higher value. Therefore the results do not seem surprising.
```{r code-oyo4}
m_e_1 <- lm(runs ~ new_onbase, data = mlb11)
m_e_2 <- lm(runs ~ new_slug, data = mlb11)
m_e_3 <- lm(runs ~ new_obs, data = mlb11)

plot(mlb11$runs ~ mlb11$new_onbase)
cor(mlb11$runs, mlb11$new_onbase)
abline(m_e_1)

plot(mlb11$runs ~ mlb11$new_slug)
cor(mlb11$runs, mlb11$new_slug)
abline(m_e_2)

plot(mlb11$runs ~ mlb11$new_obs)
cor(mlb11$runs, mlb11$new_obs)
abline(m_e_3)
```

### OYO 5
Linearity: The data does show a linear trend with a strong correlation coefficient.

Nearly Normal Residuals: The histogram of the residuals resembles a normal distribution closely.

Constant Variability: The residual plot shows constant variability, as do the points on the normal probability plot.

Independent Observations: We know from the dataset that each observation is independent.
```{r code-oyo5}
plot(m_e_3$residuals ~ mlb11$new_obs)
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
hist(m_e_3$residuals)
qqnorm(m_e_3$residuals)
qqline(m_e_3$residuals)  # adds diagonal line to the normal prob plot
summary(m_e_3)
```
