---
title: "Lab 4: The Normal Distribution"
author: "Daniel Deneau"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
```

### Exercise 1

The male heights are normally distributed, with a median around 176. While the female heights are skewed slightly towards lower heights, with a median around 173. Males have a higher variability as well.

```{r code-chunk-1}
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)

hist(mdims$hgt, probability = TRUE)
hist(fdims$hgt, probability = TRUE)
```

### Exercise 2
The normal distribution line does fit the histogram of female heights. With the lowest occuring heights reaching the ends of the distribution, and the most frequent heights in the center.
```{r code-chunk-2}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)

hist(fdims$hgt, probability = TRUE)
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

### Exercise 3
While the simulated points are close to the distribution line, a lot of the points are above or below the line. This simulated probability plot is close to the one for the real data.
```{r code-chunk-3}
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm)
qqline(sim_norm)
```

### Exercise 4
These created plots look very similar to the ones created for the simlulated data, which provides evidence that female heights are nearly normal.
```{r code-chunk-4}
qqnormsim(fdims$hgt)
qqnormsim(sim_norm)
```

### Exercise 5
Based on the multiple plots generated of both the simulated and real female weights, it does seem that female weights also come from a normal distribution. The histogram shows features close to having a perfect normal distribution such as one center and symmetry, but is skewed left. But the similarities in the normal qq plots show evidence that female weights come from a female distribution.  
```{r code-chunk-5}
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)

hist(fdims$wgt, probability = TRUE)
x <- 140:190
y <- dnorm(x = x, mean = fwgtmean, sd = fwgtsd)
lines(x = x, y = y, col = "blue")

sim_norm_wgt <- rnorm(n = length(fdims$wgt), mean = fwgtmean, sd = fwgtsd)
qqnorm(sim_norm_wgt)
qqline(sim_norm_wgt)

qqnormsim(fdims$wgt)
qqnormsim(sim_norm_wgt)

```

### Exercise 6
What are the chances of a random female being under 160 centimeters?
Using p-norm: 0.2283
Empirically: 0.1923
Difference = 0.016

What are the chances of a random female being over 50 kilograms?
Using p-norm: 0.8648
Empirically: 0.8808
Difference = 0.0360

The closest agreement was with the height variable.
```{r code-chunk-6}
pnorm(q = 160, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt < 160) / length(fdims$hgt)

1 - pnorm(q = 50, mean = fwgtmean, sd = fwgtsd)
sum(fdims$wgt > 50) / length(fdims$wgt)

```

# OYO 1
a) A
b) D
c) C
d) B

# OYO 2
Because the variables age for those two plots have patterns of sharp increases and decreases.

# OYO 3
The variable is skewed left.
```{r code-chunk-oyo3}


hist(fdims$kne.di, probability = TRUE)
```