---
title: "Lab 1: Intro to R"
author: "Daniel Deneau"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

Overall there is a positive trend in the number of girl baptisms 
with a few periods of decline.

```{r trend-girls}
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
```


### Exercise 3

There seems to be a very slight trend towards proportionally less boy baptisms throughout the years. Indicating that the gap between boy and girl baptisms decreased over time.

```{r plot-prop-boys-arbuthnot}
plot(x = arbuthnot$year, y = (arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)))
```


### Exercise 4

The present dataset is 3 columns, year, boys, and girls, and has 63 entries.
Each entry shows the year, number of boys born, and the number of girls born.

```{r dim-present}
dim(present)
```


### OYO 1

The present set has a count of 63 with a scale in the millions, while the arbuthnot set has a count of 82, with a scale only in the thousands 17. The range of time in the present data set is 62 years, which is less than the 81 years in the arbuthnot set.

```{r count-compare}
count(present)
count(arbuthnot)
```


### OYO 2

The trend of boys being born in greater proportion than girls is shown in the plot, as
the proportion does not drop below 1. However, the trend also shows that that proportion is has been decreasing linearly over time.

```{r plot-prop-boys-present}
plot(x = present$year, y = (present$boys / present$girls))
```


### OYO 3

The maximum number of total births, 4268326, was in 1961.

```{r find-max-total}
max((present$boys + present$girls))
which.max((present$boys + present$girls))
present[22,] 
```

